Quanta concentration and Transmission risk

### 2024-03-21 analysis/patient-tracking.Rmd
Results and prior vis updates

```{r}
track_qual <- df %>%
  group_by(date, patient_id) %>%
  slice(1) %>%
  ungroup() %>%
  dplyr::select(tracking_end) %>%
  unlist() %>%
  table()

sum(track_qual[names(track_qual) != "Possible HCW"])
track_qual_str <- paste0(
  names(track_qual),
  ": ",
  track_qual,
  " (",
  round(track_qual / sum(track_qual) * 100),
  "%)"
)
track_qual_str

df %>%
  filter(tracking_end != "Possible HCW") %>%
  group_by(date, patient_id) %>%
  slice(1) %>%
  ungroup() %>%
  group_by(tracking_end) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  mutate(prop = round(n / sum(n) * 100))
```

>>>>

```{r}
track_qual <- df_filt %>%
  group_by(date, patient_id) %>%
  slice(1) %>%
  ungroup() %>%
  dplyr::select(tracking_end) %>%
  unlist() %>%
  table()

track_qual_str <- paste0(
  names(track_qual),
  ": ",
  track_qual,
  " (",
  round(track_qual / sum(track_qual) * 100),
  "%)"
)
track_qual_str

df_filt %>%
  group_by(date, patient_id, tracking_end) %>%
  summarise(time_in_clinic = n() / 60) %>%
  ungroup() %>%
  group_by(tracking_end) %>%
  tidybayes::mean_qi(time_in_clinic, .width = .95) %>%
  ungroup()
```

### 2024-03-21 analysis/qctr.Rmd
Results and prior vis updates

```{r}
# data
mth <- 10
dy <- 25
sim <- 1
param_1025 <- readRDS(
  paste0(
    "../simulations/", "/", "sim-param-",
    2021, "-", mth, "-", dy, ".rds"
  )
)
track_1025 <- readRDS(
  paste0(
    "../simulations/", "/", "tracking-df-",
    250, "px-", 2021, "-", mth, "-", dy, ".rds"
  )
)
track_undiag <- track_1025 %>%
  dplyr::filter(tracking_end != "Possible HCW")
track_diag <- dplyr::filter(track_1025, is_infectious)
track_undiag <- dplyr::filter(
  track_undiag, !is_infectious | is.na(is_infectious)
)
track_tb_all <- rbind(
  track_diag,
  filter(
    track_undiag,
    patient_id %in% param_1025[[paste0("undiagTB", "_", "du")]][[sim]]
  )
)
aer_df <- readRDS("../data-clean/environmental/air-change-rate.rds") %>%
  filter(
    month(date) == as.numeric(mth),
    day(date) == as.numeric(dy)
  )

@@

```{r}
main_files <- c(
  list.files("simulations/main/10-13/", full.names = T),
  list.files("simulations/main/10-15/", full.names = T),
  list.files("simulations/main/10-25/", full.names = T),
  list.files("simulations/main/11-04/", full.names = T),
  list.files("simulations/main/11-05/", full.names = T)
)
```

@@

# example translation
source("utils/trans_risk.r")
IR_sit <- .5 * IR("sitting", "male") + .5 * IR("sitting", "female")
WR_vol <- convert_dist(cellSize)^2 * dimWR[3]
round((1 - exp(-max(main_qc_wr_mo$value) * 600 * IR_sit / WR_vol)) * 100, 0)
round((1 - exp(-max(main_qc_wr_af$value) * 600 * IR_sit / WR_vol)) * 100, 0)

@@

# plot
mean_quanta_pl <- plot_spatial(
  ggplot(mapping = aes(x = long, y = lat, group = group, fill = value)) +
    facet_wrap(~daytime) +
    geom_polygon(
      data = left_join(
        fortify(waiting_room),
        filter(main_qc, room == "Waiting room"),
        by = c("id")
      )
    ) +
    geom_polygon(
      data = left_join(
        fortify(corridor),
        filter(main_qc, room == "Corridor"),
        by = c("id")
      )
    ) +
    geom_polygon(
      data = left_join(fortify(tb_room),
        filter(main_qc, room == "TB room"),
        by = c("id")
      )
    ) +
    labs(
      fill = expression("Mean quanta concentration (quanta/m"^3 * " x 10e"^-3 * " per sec)"),
      title = "a"
    ) +
    scale_fill_gradientn(
      colours = RColorBrewer::brewer.pal(n = 9, name = "YlOrRd"),
      labels = function(x) round(1e3 * x / (0.25^2 * 3), 1)
    ),
  8
) +
  theme(
    legend.position = "bottom",
    legend.direction = "horizontal",
    legend.key.height = unit(0.2, "cm"),
    legend.key.width = unit(2, "cm")
  ) +
  guides(fill = guide_colorbar(title.position = "top"))

@@

  average_tr <- tr %>%
    mutate(patient_id = paste(mthdy, patient_id, sep = "-")) %>%
    group_by(patient_id) %>%
    summarize(
      P_mean = mean(P),
      P_med = median(P)
    ) %>%
    ungroup() %>%
    left_join(track %>%
      mutate(
        mth = lubridate::month(time),
        dy = lubridate::day(time)
      ) %>%
      group_by(patient_id, mth, dy) %>%
      slice(1) %>%
      ungroup() %>%
      mutate(dy = ifelse(nchar(dy) == 1, paste0("0", dy), dy)) %>%
      mutate(patient_id = paste(
        mth, dy,
        patient_id,
        sep = "-"
      )) %>%
      dplyr::select(patient_id, tracking_end)) %>%
    filter(tracking_end != "Possible HCW") %>%
    dplyr::select(-tracking_end) %>%
    reshape2::melt("patient_id") %>%
    mutate(
      val_col = ifelse(value < .05, "Low",
        ifelse(value < .1, "Medium", "High")
      ),
      val_col = factor(val_col, levels = c("Low", "Medium", "High")),
      variable = ifelse(
        grepl("mean", variable),
        "Mean risk of infection",
        "Medium risk of infection"
      ),
      variable = factor(
        variable,
        levels = paste(c("Mean", "Medium"), "risk of infection")
      )
    )

@@

# annotation
annotate_average_trans_risk <- function(tr) {
  roi_annon <- tr %>%
    group_by(variable) %>%
    summarize(
      mean_P = mean(value),
      lower_P = quantile(value, .025),
      upper_P = quantile(value, .975)
    ) %>%
    ungroup() %>%
    mutate(across(c(mean_P, lower_P, upper_P), ~ round_k(100 * .x, 1)),
      lab = paste0(
        mean_P, "%",
        " (95%-CrI ", lower_P, "% – ", upper_P, "%)"
      )
    )

@@

# plot
roi_pl <- ggplot(main_tr, aes(x = value)) +
  geom_histogram(
    mapping = aes(fill = val_col),
    binwidth = .05,
    boundary = 1
  ) +
  geom_text(
    data = main_annon,
    mapping = aes(x = Inf, y = Inf, label = lab),
    size = 8 / cm(1),
    hjust = 1, vjust = 2
  ) +
  facet_wrap(~variable) +
  scale_fill_manual(values = wes_palette("Darjeeling1")[c(2, 3, 1)]) +
  coord_cartesian(expand = FALSE) +
  scale_x_continuous(
    breaks = c(0, .05, 0.1, seq(0.2, .5, .1)),
    labels = function(x) x * 100
  ) +
  scale_y_continuous(
    labels = scales::comma
  ) +
  labs(
    x = "Risk of infection (%)",
    y = "Number of clinical attendees",
    title = "b"
  ) +
  theme_custom() +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    legend.key.height = unit(.6, "cm"),
    legend.key.width = unit(.4, "cm"),
    panel.spacing = unit(2, "cm")
  )

@@

assoc_df <- track_ft %>%
  left_join(
    aer %>%
      filter(room == "Waiting room") %>%
      dplyr::select(date, daytime, aer)
  ) %>%
  mutate(patient_id = paste(
    lubridate::month(date),
    lubridate::day(date),
    patient_id,
    sep = "-"
  )) %>%
  left_join(
    reshape2::dcast(main_tr, patient_id ~ variable) %>%
      dplyr::select(patient_id, `Mean risk of infection`) %>%
      rename(mean_risk = `Mean risk of infection`)
  )

# model
amod <- stan_betareg(mean_risk ~ aer + time_in_clinic + n_cc + cc_time,
  data = assoc_df %>% mutate(
    n_cc = n_cc + 1, cc_time = cc_time + 1 / 60,
    across(c(aer, time_in_clinic, n_cc, cc_time), log2)
  ),
  seed = 12345, cores = 4
)

amod_est <- spread_draws(amod, aer, time_in_clinic, n_cc, cc_time) %>%
  reshape2::melt(c(".chain", ".iteration", ".draw")) %>%
  mutate(value = exp(value))

amod_est %>%
  group_by(variable) %>%
  summarize(
    lower = quantile(value, .025),
    median = median(value),
    upper = quantile(value, .975)
  ) %>%
  mutate(across(c(lower, median, upper), ~ round_k(.x, 2)))

amod_est_pl <- ggplot(amod_est, aes(y = variable, x = value)) +
  geom_vline(aes(xintercept = 1), linetype = "dotted", color = "red") +
  stat_interval() +
  stat_summary(
    geom = "point",
    fun = median,
    size = 2,
    shape = 21,
    fill = "white"
  ) +
  labs(
    x = "Odds ratio of the modeled risk of infection for a doubling in the predictor",
    title = "c",
    color = "Credible interval"
  ) +
  scale_y_discrete(
    labels = c(
      "Air change rate",
      "Time in clinic",
      "Number of close contacts",
      "Time in clocse contact"
    )
  ) +
  scale_color_brewer(
    labels = paste0(c(95, 80, 50), "%")
  ) +
  theme_custom() +
  theme(axis.title.y = element_blank())

@@

```{r main-figure}
main_fig <- grid.arrange(
  mean_quanta_pl,
  roi_pl,
  amod_est_pl,
  ncol = 1,
  heights = c(8, 8, 4)
)

ggsave(
  main_fig,
  filename = "results/modeling/main-figure.png",
  width = 16 / cm(1),
  height = 20 / cm(1)
)
```

>>>>

```{r illustration}
# data
mth <- 10
dy <- 25
sim <- 1
param_1025 <- readRDS(
  paste0(
    "../simulations/", "/", "sim-param-",
    2021, "-", mth, "-", dy, ".rds"
  )
)
track_1025 <- readRDS(
  paste0(
    "../simulations/", "/", "tracking-df-",
    250, "px-", 2021, "-", mth, "-", dy, ".rds"
  )
)
track_undiag <- track_1025 %>%
  dplyr::filter(tracking_end != "Possible HCW")
track_diag <- dplyr::filter(track_1025, is_infectious)
track_undiag <- dplyr::filter(
  track_undiag, !is_infectious | is.na(is_infectious)
)
track_tb_all <- rbind(
  track_diag,
  filter(
    track_undiag,
    patient_id %in% param_1025[[paste0("undiagTB", "_", "du")]][[sim]]
  )
)
aer_df <- readRDS("../data-clean/environmental/air-change-rate.rds") %>%
  filter(
    month(date) == as.numeric(mth),
    day(date) == as.numeric(dy)
  )

@@

```{r main-files}
main_files <- list.files("simulations/main/",
  full.names = TRUE, pattern = "*.rds$",
  recursive = TRUE
)
```

@@

# example translation
source("utils/trans_risk.r")
IR_sit <- (.5 * IR("sitting", "male") + .5 * IR("sitting", "female")) / 3600
WR_vol <- convert_dist(cellSize)^2 * dimWR[3]
round((1 - exp(-max(main_qc_wr_mo$value) * 600 * IR_sit / WR_vol)) * 100, 3)
round((1 - exp(-max(main_qc_wr_af$value) * 600 * IR_sit / WR_vol)) * 100, 3)

@@

# plot
mean_quanta_pl <- plot_spatial(
  ggplot(mapping = aes(x = long, y = lat, group = group, fill = value)) +
    facet_wrap(~daytime) +
    geom_polygon(
      data = left_join(
        fortify(waiting_room),
        filter(main_qc, room == "Waiting room"),
        by = c("id")
      )
    ) +
    geom_polygon(
      data = left_join(
        fortify(corridor),
        filter(main_qc, room == "Corridor"),
        by = c("id")
      )
    ) +
    geom_polygon(
      data = left_join(fortify(tb_room),
        filter(main_qc, room == "TB room"),
        by = c("id")
      )
    ) +
    labs(
      fill = expression("Mean quanta concentration (quanta m"^-3 * " " * "x 10e"^-3 * ")"),
      title = "a"
    ) +
    scale_fill_gradientn(
      colours = RColorBrewer::brewer.pal(n = 9, name = "YlOrRd"),
      labels = function(x) round(1e3 * x / (0.25^2 * 3), 1)
    ),
  8
) +
  theme(
    legend.position = "bottom",
    legend.direction = "horizontal",
    legend.key.height = unit(0.2, "cm"),
    legend.key.width = unit(2, "cm")
  ) +
  guides(fill = guide_colorbar(title.position = "top"))

@@

  average_tr <- tr %>%
    mutate(patient_id = paste(mthdy, patient_id, sep = "-")) %>%
    group_by(patient_id) %>%
    summarize(
      P_mean = mean(P),
      P_med = median(P)
    ) %>%
    ungroup() %>%
    left_join(track %>%
      mutate(
        mth = lubridate::month(time),
        dy = lubridate::day(time)
      ) %>%
      group_by(patient_id, mth, dy) %>%
      slice(1) %>%
      ungroup() %>%
      mutate(dy = ifelse(nchar(dy) == 1, paste0("0", dy), dy)) %>%
      mutate(patient_id = paste(
        mth, dy,
        patient_id,
        sep = "-"
      )) %>%
      dplyr::select(patient_id, tracking_end)) %>%
    filter(tracking_end != "Possible HCW") %>%
    dplyr::select(-tracking_end) %>%
    reshape2::melt("patient_id") %>%
    mutate(
      val_col = ifelse(value < .001, "Low",
        ifelse(value < .01, "Medium", "High")
      ),
      val_col = factor(val_col, levels = c("Low", "Medium", "High")),
      val_ann = 1 - (1 - value)^12,
      variable = ifelse(
        grepl("mean", variable),
        "Mean risk of infection",
        "Medium risk of infection"
      ),
      variable = factor(
        variable,
        levels = paste(c("Mean", "Medium"), "risk of infection")
      )
    )

@@

# annotation
annotate_average_trans_risk <- function(tr) {
  roi_annon <- tr %>%
    group_by(variable) %>%
    summarize(
      mean_P = mean(value),
      lower_P = quantile(value, .025),
      upper_P = quantile(value, .975),
      mean_P_ann = mean(val_ann),
      lower_P_ann = quantile(val_ann, .025),
      upper_P_ann = quantile(val_ann, .975)
    ) %>%
    ungroup() %>%
    mutate(
      across(c(mean_P, lower_P, upper_P), ~ round_k(100 * .x, 3)),
      lab = paste0(
        mean_P, "%",
        " (95%-CrI ", lower_P, "% – ", upper_P, "%)"
      ),
      across(c(mean_P_ann, lower_P_ann, upper_P_ann), ~ round_k(100 * .x, 2)),
      lab_ann = paste0(
        mean_P_ann, "%",
        " (95%-CrI ", lower_P_ann, "% – ", upper_P_ann, "%)"
      )
    )

@@

main_annon

# plot
roi_pl <- ggplot(
  data = main_tr %>% filter(variable == "Mean risk of infection"),
  mapping = aes(x = value)
) +
  geom_histogram(
    mapping = aes(fill = val_col),
    binwidth = .0005,
    boundary = 1
  ) +
  geom_text(
    data = main_annon %>% filter(variable == "Mean risk of infection"),
    mapping = aes(x = Inf, y = Inf, label = lab),
    size = 8 / cm(1),
    hjust = 1, vjust = 2
  ) +
  scale_fill_manual(values = wes_palette("Darjeeling1")[c(2, 3, 1)]) +
  coord_cartesian(expand = FALSE) +
  scale_x_continuous(
    labels = function(x) x * 100
  ) +
  scale_y_continuous(
    labels = scales::comma
  ) +
  labs(
    x = "Risk of infection (%)",
    y = "Number of clinical attendees",
    title = "b"
  ) +
  theme_custom() +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    legend.key.height = unit(.6, "cm"),
    legend.key.width = unit(.4, "cm"),
    panel.spacing = unit(2, "cm")
  )

@@

assoc_df <- track_ft %>%
  # left_join(
  #   aer %>%
  #     filter(room == "Waiting room") %>%
  #     dplyr::select(date, daytime, aer)
  # ) %>%
  mutate(patient_id = paste(
    lubridate::month(date),
    lubridate::day(date),
    patient_id,
    sep = "-"
  )) %>%
  left_join(
    reshape2::dcast(main_tr, patient_id ~ variable) %>%
      dplyr::select(patient_id, `Mean risk of infection`) %>%
      rename(mean_risk = `Mean risk of infection`)
  )

# model
amod <- stan_betareg(mean_risk ~ time_in_clinic + n_cc + cc_time,
  data = assoc_df %>% mutate(
    n_cc = n_cc + 1, cc_time = cc_time + 1 / 60,
    across(c(time_in_clinic, n_cc, cc_time), log2)
  ),
  seed = 12345, cores = 4
)

amod_est <- spread_draws(amod, time_in_clinic, n_cc, cc_time) %>%
  reshape2::melt(c(".chain", ".iteration", ".draw")) %>%
  mutate(value = exp(value))

amod_est %>%
  group_by(variable) %>%
  summarize(
    lower = quantile(value, .025),
    mean = mean(value),
    upper = quantile(value, .975)
  ) %>%
  mutate(across(c(lower, mean, upper), ~ round_k(.x, 2)))

amod_est_pl <- ggplot(amod_est, aes(y = variable, x = value)) +
  geom_vline(aes(xintercept = 1), linetype = "dotted", color = "red") +
  stat_interval(point_interval = "mean_qi") +
  stat_summary(
    geom = "point",
    fun = mean,
    size = 2,
    shape = 21,
    fill = "white"
  ) +
  labs(
    x = "OR of modeled risk if 2 x predictor",
    title = "c",
    color = "CrI"
  ) +
  scale_y_discrete(
    labels = c(
      # "Air change rate",
      "Time in clinic",
      "No. close contacts",
      "Time close contact"
    )
  ) +
  scale_color_brewer(
    labels = paste0(c(95, 80, 50), "%")
  ) +
  theme_custom() +
  theme(
    axis.title.y = element_blank(),
    legend.position = "bottom"
  )

@@

```{r main-figure}
main_fig <- grid.arrange(
  mean_quanta_pl,
  grid.arrange(roi_pl, amod_est_pl, ncol = 2),
  ncol = 1,
  heights = c(8, 8)
)

ggsave(
  main_fig,
  filename = "results/modeling/main-figure.png",
  width = 16 / cm(1),
  height = 16 / cm(1)
)
```
